{\bf Practical Concurrency Testing}

### 2018-10-14 thesis/defense/snakefight.tex
update snaekfight after first practice

\newcommand\hilight[2]{\color{#1}#2\color{black}}
\definecolor{darkorange}{RGB}{192,128,0}
\definecolor{darkblue}{RGB}{0,97,192}
\definecolor{olivegreen}{RGB}{0,128,0}
\definecolor{lavender}{RGB}{102,65,208} % (?? i don't know what happened here anymore but it looks good)
\definecolor{seafoam}{RGB}{56,158,68} % * 2/3
\definecolor{salmon}{RGB}{208,89,74} % * 7/8
\definecolor{goldish}{RGB}{128,96,32}
\definecolor{gray}{RGB}{128,128,128}

\begin{document}
\renewcommand{\inserttotalframenumber}{39}
\normalem
\begin{frame}
	\titlepage
\end{frame}

@@

\begin{frame}{Concurrency!}
	\begin{center}
	\begin{tabular}{ll}
		Initially \texttt{x = 0;} \\
		\\
		{\bf Thread 1} & {\bf Thread 2} \\
		\\
		\texttt{\hilight{darkorange}{mutex\_lock}(m);}      & \texttt{atomic\_xadd(\&x, 1);} \\
		\texttt{x++;}				   & \texttt{\hilight{olivegreen}{yield}();} \\
		\texttt{\hilight{darkblue}{mutex\_unlock}(m);}      & \texttt{atomic\_xadd(\&x, 1);} \\
		\texttt{assert(x >= 1);}			& \texttt{assert(x >= 2);} \\
		\\
		\\
		\\
		\\
		\\
	\end{tabular}
	\end{center}
\end{frame}
\begin{frame}{Concurrency!}
	\begin{center}
	\begin{tabular}{ll}
		Initially \texttt{x = 0;} \\
		\\
		{\bf Thread 1} & {\bf Thread 2} \\
		\\
		\texttt{\hilight{darkorange}{mutex\_lock}(m);} \\
		\texttt{int tmp = x;} \\
								& \texttt{atomic\_xadd(\&x, 1);} \\
								& \texttt{\hilight{olivegreen}{yield}();} \\
								& \texttt{atomic\_xadd(\&x, 1);} \\
		\texttt{x = tmp + 1;} \\
								& \texttt{\hilight{red}{assert(x >= 2);}} \\
		\texttt{\hilight{darkblue}{mutex\_unlock}(m);} \\
		\texttt{assert(x >= 1);}
	\end{tabular}
	\end{center}
\end{frame}

@@

\begin{frame}{Challenges} %\& Existing Coping Techniques}
	Unpredictable state space explosion
	\begin{itemize}
		\item Equivalence reductions: DPOR {\em [Flanagan '05]}, MCR {\em [Huang '15]}
		\item Heuristic search ordering: ICB {\em [Musuvathi '08]}, PCT {\em [Burckhardt '10]}
		\item There will always be a {\bf bigger test case}...
	\end{itemize}
	\pause
	\linegap

	Too much manual intervention %for non-expert use
	\begin{itemize}
		\item Manual kernel annotations: Landslide {\em [Blum '12]} {\tiny that's me!}
			% TODO: smth about choosing pps?
		\item State space size estimation {\em [Simsa '12]}
			% TODO change this bullet pt to be more explicitly about 
		\item Widespread adoption requires {\bf fully automated instrumentation}...
	\end{itemize}
	\pause
	\linegap

	Sources of nondeterminism
	\begin{itemize}
		\item Timer-driven schedule nondeterminism (all prior work)
		\item Weak memory nondeterminism {\em [Zhang '15]}
		\item Event-driven applications {\em [Jensen '15]}
		\item {\bf Hardware transactional memory}...
	\end{itemize}
\end{frame}

@@

\begin{frame}{Thesis Statement}
	% data race pps (as well as txn equivalence)
	Combining both theoretically-founded automatic reduction techniques
	% iterative deepening
	% (say smth like: "landslide also has a wealth of other user-tunable heuristics
	% for expert users, which you can read about in the document; for this talk i will focus just on the CPU budget thing")
	% XXX: need to prepare for the question -- "what evidence do you have that users actually want this cpu budget stuff",
	% ",or, that it helps vs just running something unlimited vs a normal icb test"
	% (well, for the 2nd half of the Q, i can talk about how ID outperforms ICB on verficiations)
	and user-informed heuristic ones,
	stateless model checking can sufficiently mitigate exponential explosion to be
	% education
	a practical testing technique for inexperienced users
	% htm
	% TODO: this last part you just added for this talk; does the thesis version of the statmeent need to include too?
	and real-world programs alike.
\end{frame}

\begin{frame}{Thesis Statement}
	% data race pps (as well as txn equivalence)
	\hilight{gray}{Combining both theoretically-founded automatic reduction techniques
	% iterative deepening
	% (say smth like: "landslide also has a wealth of other user-tunable heuristics
	% for expert users, which you can read about in the document; for this talk i will focus just on the CPU budget thing")
	% XXX: need to prepare for the question -- "what evidence do you have that users actually want this cpu budget stuff",
	% ",or, that it helps vs just running something unlimited vs a normal icb test"
	% (well, for the 2nd half of the Q, i can talk about how ID outperforms ICB on verficiations)
	and user-informed heuristic ones, stateless model checking...}
	\begin{itemize}
		\item ...can sufficiently mitigate exponential explosion...
		% education
		\item ...to be a practical testing technique for inexperienced users...
		% htm
		% TODO: this last part you just added for this talk; does the thesis version of the statmeent need to include too?
		\item ...and real-world programs alike.
	\end{itemize}
\end{frame}

\begin{frame}{Thesis Statement}
	The Iterative Deepening algorithm %combines theory and heuristics to
	practically debugs and verifies many exponential state spaces.
	\vspace{4.54em}

@@

	New reduction techniques for transactional memory keep MC up-to-date
	with the latest real-world concurrency techniques.
	\vspace{4.54em}
\end{frame}
\begin{frame}{Outline}
	\hilight{gray}{The Iterative Deepening algorithm %combines theory and heuristics to
	practically debugs and verifies many exponential state spaces.} \\
	%\hilight{gray}{The Iterative Deepening algorithm makes MC appropriate for both educational and real-world programs.} \\
	\begin{itemize}
		\item {\bf Quicksand and Iterative Deepening}
		\begin{itemize}
			\item Data-race preemption points \& algorithm overview
			\item Verification \& experimental results
		\end{itemize}
	\end{itemize}
	\hilight{gray}{Landslide's automatic instrumentation makes MC accessible to students.} \\
	%\hilight{gray}{Ad-hoc synchronization heuristics make MC accessible to student programmers.} \\
	\begin{itemize}
		\item {\bf Educational use}
		\begin{itemize}
			\item Pebbles \& Pintos OS projects
			\item Bug-finding, grades, \& surveys
		\end{itemize}
	\end{itemize}

@@

	State space of interleavings is {\em parameterized} by PPs.
	\begin{itemize}
		\item Too few PPs: Many bugs will go undetected
		\item Preempt everywhere: Completion is often infeasible % say: "for some programs"
	\end{itemize}
	\linegap

	Prior model checkers hard-code a fixed set of PPs, committing to one state space in advance.
	% say: "...then attempt to cope with the resulting exponentially sized problem, however big it may be"
\end{frame}

\begin{frame}{Example (Revisited)}
	\begin{center}
	\begin{tabular}{ll}
		Initially \texttt{x = 0;} \\
		\\
		{\bf Thread 1} & {\bf Thread 2} \\
		\\
		\texttt{\hilight{darkorange}{mutex\_lock}(m);} \\
		\texttt{int tmp = x;} \\
								& \texttt{atomic\_xadd(\&x, 1);} \\
								& \texttt{\hilight{olivegreen}{yield}();} \\
								& \texttt{atomic\_xadd(\&x, 1);} \\
		\texttt{x = tmp + 1;} \\
								& \texttt{\hilight{red}{assert(x >= 2);}} \\
		\texttt{\hilight{darkblue}{mutex\_unlock}(m);} \\
		\texttt{assert(x >= 1);}
	\end{tabular}
	\end{center}
\end{frame}

\begin{frame}{State Space (Revisited)}
	\begin{center}
		\includegraphics[width=0.96\textwidth]{../../oopsla/tree-maximal-only.pdf}
		\linegap

@@

		A {\bf data-race preemption point} is required to find the bug.
	\end{center}
\end{frame}

\begin{frame}{Data Race Analysis}
	A {\bf data race} is a pair of memory accesses between two threads, where:
	\begin{itemize}
		\item At least one of the accesses is a write
		\item The threads are not holding the same mutex
		\item One access {\em happens-before} the other
			(e.g., no \texttt{cond\_signal()} in between)
	\end{itemize}
	\pause
	\linegap

	Two flavors
	\begin{itemize}
		\item {\bf Pure HB}: Whether accesses were {\em observed simultaneously}, false-negative prone
			{\em [Lamport '78]}
		\item {\bf Limited HB}: Whether threads {\em could be reordered}, false-positive prone
			{\em [O'Callahan '03]}
	\end{itemize}
\end{frame}

\begin{frame}{Iterative Deepening (in Pictures)}
	\begin{center}
	\vspace{-0.88em}
	\includegraphics[width=0.86\textwidth]{../../oopsla/dr-jobs-1.pdf}
	\end{center}
\end{frame}
\begin{frame}{Iterative Deepening (in Pictures)}
	\begin{center}
	\vspace{-0.88em}
	\includegraphics[width=0.86\textwidth]{../../oopsla/dr-jobs-2.pdf}
	\end{center}
\end{frame}
\begin{frame}{Iterative Deepening (in Pictures)}
	\begin{center}
	\vspace{-0.88em}
	\includegraphics[width=0.86\textwidth]{../../oopsla/dr-jobs-3.pdf}
	\end{center}
\end{frame}
\begin{frame}{Iterative Deepening (in Pictures)}
	\begin{center}
	\vspace{-0.88em}
	\includegraphics[width=0.86\textwidth]{../../oopsla/dr-jobs-4.pdf}
	\end{center}
\end{frame}

@@

\begin{frame}{Total Verification}
	\textbf{Convergence theorem:} Testing maximal state space, after detecting all data races, $\equiv$ preempting everywhere!
	\begin{itemize}
		\item Insight: $\forall$ arbitrary buggy interleaving, $\exists$ equivalent interleaving % say 'under DPOR' i guess?
			using only sync/data-race PPs, \& Quicksand will eventually reach it.
	\end{itemize}
	\linegap

	Hence, Iterative Deepening provides ``best of both worlds'':
	\begin{itemize}
		\item Provides total safety guarantee for small tests
		\item Finds bugs quickly for large tests when completion is infeasible
	\end{itemize}
\end{frame}

@@

	Each tested for 10 CPU-hours with
	\begin{itemize}
		\item Quicksand, finding data-race candidates with...
			\begin{itemize}
				\item ...``Pure'' Happens-Before analysis {\em [Lamport '78]}
				\item ...``Limited'' Happens-Before analysis {\em [O'Callahan '03]}
			\end{itemize}
		\item Stand-alone Landslide using ICB {\em [Musuvathi '08]}, preempting at...
			\begin{itemize}
				\item ...synchronization APIs only {\em [Simsa '10]}
				\item ...all shared memory accesses
					%{\em [Holzmann '97]} % not spin
					{\em [Yang '08]} % inspect
			\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}{Results - Bugs found (CPU time)}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{../../oopsla/bugs-talk.pdf}
	\end{center}
\end{frame}
\begin{frame}{Results - Bugs found (wall-clock time)}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{../../oopsla/bugs-wallclock-talk.pdf}
	\end{center}
\end{frame}
\begin{frame}{Results - Full verifications}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{../../oopsla/verifs-talk.pdf}
	\end{center}
\end{frame}

@@

\begin{frame}{Setting}
	CMU (15-410, S'15-S'18), PSU (CMPSC 473, S'18)
	\begin{itemize}
		\item Thread library (``P2'')
			\begin{itemize}
				\item POSIX-like userspace thread library %(also teams of 2)
				\item Reference ``Pebbles'' kernel provided
					% say: later they implement this in p3
				\item Fixed APIs $\rightarrow$ instrumentation can be automated
				\item Still needs heuristics to handle ad-hoc synchronization
			\end{itemize}
			% FIXME: need introduce p3 here?
			% Or cut?
		\item Kernel (``P3'') - reimplement said reference kernel (CMU only)
	\end{itemize}
	\pause
	\linegap

	U. Chicago (CMSC 23000, F'17)
	\begin{itemize}
		\item Pintos %(``threads'' and ``userprog'')
		\begin{itemize}
			\item Basecode provides round-robin scheduler, synchronization, VM, loader
			\item ``threads'': Students implement priority scheduling, sleep
		%\end{itemize}
		%\linegap

@@

	Ad-hoc synchronization (e.g. {\tt while (!ready) yield();})
	\begin{itemize}
		\item Not an infinite loop bug!
		\item Heuristically ``block'' threads when 10+ {\tt yield()}s %({\tt xchg}, etc.)
		\item Existing memory analysis %(DPOR)
			on loop condition to unblock
			\begin{itemize}
					% TODO: cut this and just say it out loud?
				\item (e.g., what other thread writes to {\tt ready}?)
				% it's kinda... idk how to critique its unsoundness.
				% they self-admit to the unsoundness but like, leaving the cycle in,
				% is the wrong approach to begin with
				\item Prior work {\em [Coons '13]} likely to leave threads blocked too long
			\end{itemize}
		\item Check for false-positive blocking to avoid false-positive deadlocks
	\end{itemize}

@@

	...as well as subsequent P3 (kernel) grades
	\begin{itemize}
		\item ``Does Landslide teach students lasting lessons about concurrency?''
	\end{itemize}
\end{frame}

@@

\begin{frame}{Example}
	\begin{center}
		\begin{tabular}{l}
			\texttt{if ((status = \hilight{darkorange}{\_xbegin}()) == SUCCESS) \{} \\ %\_XBEGIN\_STARTED) \{} \\
			%\texttt{} \\
			%\texttt{} \\
			\texttt{~~~~count++;} \\
			\texttt{~~~~\hilight{darkblue}{\_xend}();} \\
			\texttt{\} else \{}\\ %if (status != \_XABORT\_EXPLICIT) \{} \\
			\texttt{~~~~\hilight{darkorange}{mutex\_lock}(\&count\_lock);} \\
			\texttt{~~~~count++;} \\
			\texttt{~~~~\hilight{darkblue}{mutex\_unlock}(\&count\_lock);} \\
			\texttt{\}} \\
		\end{tabular}
	\end{center}
	% TODO: cite orig software TM, intel HTM
\end{frame}

@@

\begin{frame}{Example bug}
	\begin{center}
		\begin{tabular}{l|l}
			{\bf Thread 1} & {\bf Thread 2} \\
			\hline
			{\tt \hilight{darkorange}{\_xbegin}() == ABORTED} \\
			{\tt \hilight{darkorange}{mutex\_lock}(...);} \\
			{\tt int tmp = count;} \\
			{\em \hilight{red}{<data-race PP>}} \\
				& {\tt \hilight{darkorange}{\_xbegin}() == SUCCESS} \\
				& {\tt count++;} \\
				& {\tt \hilight{darkblue}{\_xend}();} \\
			{\tt count = tmp + 1;} \\
			{\tt \hilight{darkblue}{mutex\_unlock}(...);} \\
		\end{tabular}
	\end{center}
\end{frame}

\begin{frame}{Example bug - Solution {\em [Dehesa-Azuara '16]}}
	\begin{center}
		\begin{tabular}{l}
			\texttt{if ((status = \hilight{darkorange}{\_xbegin}()) == SUCCESS) \{} \\ %\_XBEGIN\_STARTED) \{} \\
			%\texttt{} \\
			%\texttt{} \\
			\texttt{~~~~\hilight{red}{if (mutex\_held(\&count\_lock))}} \\
			\texttt{~~~~~~~~\hilight{red}{\_xabort();}} \\
			\texttt{~~~~count++;} \\
			\texttt{~~~~\hilight{darkblue}{\_xend}();} \\
			\texttt{\} else \{}\\ %if (status != \_XABORT\_EXPLICIT) \{} \\
			\texttt{~~~~\hilight{darkorange}{mutex\_lock}(\&count\_lock);} \\
			\texttt{~~~~count++;} \\
			\texttt{~~~~\hilight{darkblue}{mutex\_unlock}(\&count\_lock);} \\
			\texttt{\}} \\
		\end{tabular}
	\end{center}
\end{frame}

>>>>

\newcommand\hilight[2]{\color{#1}#2\color{black}}
\definecolor{darkorange}{RGB}{192,128,0}
\definecolor{darkblue}{RGB}{0,97,192}
\definecolor{olivegreen}{RGB}{0,128,0}
\definecolor{lavender}{RGB}{102,65,208} % (?? i don't know what happened here anymore but it looks good)
\definecolor{seafoam}{RGB}{56,158,68} % * 2/3
\definecolor{salmon}{RGB}{208,89,74} % * 7/8
\definecolor{pinkish}{RGB}{224,0,56} % * 7/8
\definecolor{goldish}{RGB}{128,96,32}
\definecolor{gray}{RGB}{128,128,128}

\begin{document}
\renewcommand{\inserttotalframenumber}{67}
\normalem
\begin{frame}
	\titlepage
\end{frame}

@@

\begin{frame}{Concurrency!}
	\begin{center}
	\begin{tabular}{ll}
		Initially \texttt{count = 0;} \\
		\\
		{\bf Thread 1} & {\bf Thread 2} \\
		\\
		\texttt{\hilight{darkorange}{mutex\_lock}(m);}      & \texttt{atomic\_xadd(\&count, 1);} \\
		\texttt{count++;}				   & \texttt{\hilight{olivegreen}{yield}();} \\
		\texttt{\hilight{darkblue}{mutex\_unlock}(m);}      & \texttt{atomic\_xadd(\&count, 1);} \\
		\texttt{assert(count >= 1);}			& \texttt{assert(count >= 2);} \\
		\\
		\\
		\\
		\\
		\\
	\end{tabular}
	\end{center}
\end{frame}
\begin{frame}{Concurrency!}
	\begin{center}
	\begin{tabular}{ll}
		Initially \texttt{count = 0;} \\
		\\
		{\bf Thread 1} & {\bf Thread 2} \\
		\\
		\texttt{\hilight{darkorange}{mutex\_lock}(m);} \\
		\texttt{int tmp = count;} \\
								& \texttt{atomic\_xadd(\&count, 1);} \\
								& \texttt{\hilight{olivegreen}{yield}();} \\
								& \texttt{atomic\_xadd(\&count, 1);} \\
		\texttt{count = tmp + 1;} \\
								& \texttt{\hilight{pinkish}{assert(count >= 2);}} \\
		\texttt{\hilight{darkblue}{mutex\_unlock}(m);} \\
		\texttt{assert(count >= 1);}
	\end{tabular}
	\end{center}
\end{frame}

@@

\begin{frame}{Challenges} %\& Existing Coping Techniques}
	Unpredictable state space explosion
	\begin{itemize}
		%\item Equivalence reductions: DPOR {\em [Flanagan '05]}, MCR {\em [Huang '15]}
		\item Dynamic Partial Order Reduction (DPOR) {\em [Flanagan '05]}
			\begin{itemize}
				\item identifies \& skips equivalent interleavings
			\end{itemize}
		%\item Heuristic search ordering: ICB {\em [Musuvathi '08]}, PCT {\em [Burckhardt '10]}
		\item Iterative Context Bounding (ICB) {\em [Musuvathi '08]}
			\begin{itemize}
				\item finds bugs faster by prioritizing interleavings with fewer preemptions
			\end{itemize}
			\pause
		\item There will always be a {\bf bigger test case}...
	\end{itemize}
	\pause
	\linegap

	Too much manual intervention %for non-expert use
	\begin{itemize}
		\item State space size estimation {\em [Simsa '12]}
			% TODO: i just switched the order of these; see how that feels
		\item Manual kernel annotations: Landslide {\em [Blum '12]} {\tiny that's me!}
			\pause
		\item Widespread adoption requires {\bf fully automated instrumentation}...
	\end{itemize}
	\pause
	\linegap

	Sources of nondeterminism
	\begin{itemize}
		\item Timer-driven schedule nondeterminism (all prior work)
		\item Weak memory {\em [Zhang '15]},
			event-driven {\em [Jensen '15]}
			\pause
		\item {\bf Hardware transactional memory}...
	\end{itemize}
\end{frame}

@@

\begin{frame}{Thesis Statement}
	% data race pps (as well as txn equivalence)
	Combining %both
	theoretically-founded automatic reduction techniques
	% iterative deepening
	% (say smth like: "landslide also has a wealth of other user-tunable heuristics
	% for expert users, which you can read about in the document; for this talk i will focus just on the CPU budget thing")
	% XXX: need to prepare for the question -- "what evidence do you have that users actually want this cpu budget stuff",
	% ",or, that it helps vs just running something unlimited vs a normal icb test"
	% (well, for the 2nd half of the Q, i can talk about how ID outperforms ICB on verficiations)
	and user-informed heuristic ones,
	stateless model checking can sufficiently mitigate exponential explosion to be
	% education
	a practical testing technique for inexperienced users
	% htm
	% TODO: this last part you just added for this talk; does the thesis version of the statmeent need to include too?
	and real-world programs alike.
\end{frame}

\begin{frame}{Thesis Statement}
	% data race pps (as well as txn equivalence)
	\hilight{gray}{Combining %both
	theoretically-founded automatic reduction techniques
	% iterative deepening
	% (say smth like: "landslide also has a wealth of other user-tunable heuristics
	% for expert users, which you can read about in the document; for this talk i will focus just on the CPU budget thing")
	% XXX: need to prepare for the question -- "what evidence do you have that users actually want this cpu budget stuff",
	% ",or, that it helps vs just running something unlimited vs a normal icb test"
	% (well, for the 2nd half of the Q, i can talk about how ID outperforms ICB on verficiations)
	and user-informed heuristic ones, stateless model checking...}
	\begin{itemize}
		\item ...can sufficiently mitigate exponential explosion...
		% education
		\item ...to be a practical testing technique for inexperienced users...
		% htm
		% TODO: this last part you just added for this talk; does the thesis version of the statmeent need to include too?
		\item ...and real-world programs alike.
	\end{itemize}
\end{frame}

\begin{frame}{Thesis Statement}
	The Quicksand algorithm %combines theory and heuristics to
	practically debugs and verifies many exponential state spaces.
	\vspace{4.54em}

@@

	New reduction techniques for transactional memory keep MC up-to-date
	with the latest real-world concurrency techniques.
	\vspace{4.54em}
\end{frame}
\begin{frame}{Outline}
	\hilight{gray}{The Quicksand algorithm %combines theory and heuristics to
	practically debugs and verifies many exponential state spaces.} \\
	%\hilight{gray}{The Quicksandalgorithm makes MC appropriate for both educational and real-world programs.} \\
	\begin{itemize}
		\item {\bf Quicksand}
		\begin{itemize}
			\item Data-race preemption points \& algorithm overview
			\item Verification \& experimental results
		\end{itemize}
	\end{itemize}
	\hilight{gray}{Landslide's automatic instrumentation makes MC accessible to students.} \\
	%\hilight{gray}{Ad-hoc synchronization heuristics make MC accessible to student programmers.} \\
	\begin{itemize}
		\item {\bf Educational use}
		\begin{itemize}
			\item Pebbles \& Pintos OS projects
			\item Bug-finding, grades, \& surveys
		\end{itemize}
	\end{itemize}

@@

	State space of interleavings is {\em parameterized} by PPs.
	\begin{itemize}
		\item Too few PPs: Many bugs will go undetected
		\item Preempt everywhere: Completion is often infeasible % say: "for some programs"
	\end{itemize}
\end{frame}

\begin{frame}{Example (Revisited)}
	\begin{center}
	\begin{tabular}{ll}
		Initially \texttt{count = 0;} \\
		\\
		{\bf Thread 1} & {\bf Thread 2} \\
		\\
		\texttt{\hilight{darkorange}{mutex\_lock}(m);} \\
		\texttt{int tmp = count;} \\
								& \texttt{atomic\_xadd(\&count, 1);} \\
								& \texttt{\hilight{olivegreen}{yield}();} \\
								& \texttt{atomic\_xadd(\&count, 1);} \\
		\texttt{count = tmp + 1;} \\
								& \texttt{\hilight{pinkish}{assert(count >= 2);}} \\
		\texttt{\hilight{darkblue}{mutex\_unlock}(m);} \\
		\texttt{assert(count >= 1);}
	\end{tabular}
	\end{center}
\end{frame}

\begin{frame}{State Space (Revisited)}
	\begin{center}
		\includegraphics[width=0.96\textwidth]{../../oopsla/tree-maximal-only.pdf}
	\end{center}
	\linegap

@@

	A {\bf data-race preemption point} is required to find the bug.
\end{frame}

\begin{frame}{Data Race Analysis}
	A {\bf data race} is a pair of memory accesses between two threads, where:
	\begin{itemize}
		\item At least one of the accesses is a write
		\item The threads are not holding the same mutex
		\item One access {\em happens-before} the other
			%(e.g., no \texttt{cond\_signal()} in between)
			(e.g., no \texttt{cond\_wait()})
	\end{itemize}
	\pause
	\linegap

	%Prior model checkers hard-code a fixed set of PPs, committing to one state space in advance.
	Prior model checkers commit to a fixed set of PPs in advance.
	\begin{itemize}
		\item ``Preempt on {\tt mutex\_lock()}, {\tt mutex\_unlock()} only''
		\item ``Preempt on every single memory access''
	\end{itemize}
	\linegap

	Insight: Decide where to preempt on-the-fly
	% say: "...then attempt to cope with the resulting exponentially sized problem, however big it may be"

	%Two flavors
	%\begin{itemize}
	%	\item {\bf Pure HB}: Whether accesses were {\em observed simultaneously}, false-negative prone
	%		{\em [Lamport '78]}
	%	\item {\bf Limited HB}: Whether threads {\em could be reordered}, false-positive prone
	%		{\em [O'Callahan '03]}
	%\end{itemize}
\end{frame}

\begin{frame}{Quicksand (in Pictures)}
	\begin{center}
	\vspace{-0.88em}
	\includegraphics[width=0.86\textwidth]{../../oopsla/dr-jobs-1.pdf}
	\end{center}
\end{frame}
\begin{frame}{Quicksand (in Pictures)}
	\begin{center}
	\vspace{-0.88em}
	\includegraphics[width=0.86\textwidth]{../../oopsla/dr-jobs-2.pdf}
	\end{center}
\end{frame}
\begin{frame}{Quicksand (in Pictures)}
	\begin{center}
	\vspace{-0.88em}
	\includegraphics[width=0.86\textwidth]{../../oopsla/dr-jobs-3.pdf}
	\end{center}
\end{frame}
\begin{frame}{Quicksand (in Pictures)}
	\begin{center}
	\vspace{-0.88em}
	\includegraphics[width=0.86\textwidth]{../../oopsla/dr-jobs-4.pdf}
	\end{center}
\end{frame}
\begin{frame}{Quicksand (in Pictures)}
	\begin{center}
	\vspace{-0.88em}
	\includegraphics[width=0.86\textwidth]{dr-jobs-maximal.pdf}
	\end{center}
\end{frame}

@@

\begin{frame}{Total Verification}
	\textbf{Convergence theorem:} Testing maximal state space, after detecting all data races, $\equiv$ preempting everywhere!
	%\begin{itemize}
	%	\item Insight: $\forall$ arbitrary buggy interleaving, $\exists$ equivalent interleaving % say 'under DPOR' i guess?
	%		using only sync/data-race PPs, \& Quicksand will eventually reach it.
	%\end{itemize}
	\linegap

	Hence, Quicksand provides ``best of both worlds'':
	\begin{itemize}
		\item Provides total safety guarantee for small tests
		\item Finds bugs quickly for large tests when completion is infeasible
	\end{itemize}
\end{frame}

@@

	Each tested for 10 CPU-hours with
	\begin{itemize}
		\item Quicksand % parallelized 10-fold
			%, finding data-race candidates with...
			%\begin{itemize}
			%	\item ...``Pure'' Happens-Before analysis {\em [Lamport '78]}
			%	\item ...``Limited'' Happens-Before analysis {\em [O'Callahan '03]}
			%\end{itemize}
		\item Stand-alone Landslide using ICB {\em [Musuvathi '08]}, preempting at...
			\begin{itemize}
				\item ...synchronization APIs only {\em [Simsa '10]}
				\item ...all shared memory accesses
					%{\em [Holzmann '97]} % not spin
					{\em [Yang '08]} % inspect
			\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}{Results - Bugs found (CPU time)}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{../quicksand-bugs-talkver.pdf}
	\end{center}
\end{frame}
\begin{frame}{Results - Bugs found (wall-clock time)}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{../quicksand-bugs-wallclock-talkver.pdf}
	\end{center}
\end{frame}
\begin{frame}{Results - Full verifications}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{../quicksand-verifs-talkver.pdf}
	\end{center}
\end{frame}

@@

\begin{frame}{Setting}
	%CMU (15-410, S'15-S'18), PSU (CMPSC 473, S'18)
	CMU (15-410), PSU (CMPSC 473)
	\begin{itemize}
		\item Thread library (``P2'')
			\begin{itemize}
				\item POSIX-like userspace thread library %(also teams of 2)
				\item Reference ``Pebbles'' kernel provided
					% say: later they implement this in p3
				\item Fixed APIs $\rightarrow$ instrumentation can be automated
				\item Still needs heuristics to handle ad-hoc synchronization
			\end{itemize}
			% FIXME: need introduce p3 here?
			% Or cut?
		\item Kernel (``P3'') - reimplement said reference kernel (CMU only)
	\end{itemize}
	\pause
	\linegap

	%U. Chicago (CMSC 23000, F'17)
	U. Chicago (CMSC 23000)
	\begin{itemize}
		\item Pintos %(``threads'' and ``userprog'')
		\begin{itemize}
			\item Basecode provides round-robin scheduler, synchronization, VM, loader
			\item ``threads'': Students implement priority scheduling, sleep
		%\end{itemize}
		%\linegap

@@

	Ad-hoc synchronization (e.g. {\tt while (!ready) yield();})
	\begin{itemize}
		\item Not an infinite loop bug!
		\item Heuristically ``block'' threads when 10+ {\tt yield()}s %({\tt xchg}, etc.)
		\item Existing memory analysis %(DPOR)
			on loop condition to unblock
			\begin{itemize}
					% TODO: cut this and just say it out loud?
				\item (e.g., what other thread writes to {\tt ready}?)
				% it's kinda... idk how to critique its unsoundness.
				% they self-admit to the unsoundness but like, leaving the cycle in,
				% is the wrong approach to begin with
					% TODO: move this to bonus slides, save a min of expl
				%\item Prior work {\em [Coons '13]} likely to leave threads blocked too long
			\end{itemize}
			\pause
		\item ``Ok, what if I just {\tt yield()} 10 times because I feel like it?''
			\begin{itemize}
				\item If deadlock suspected, try waking up to avoid false positives
			\end{itemize}
		%\item Check for false-positive blocking to avoid false-positive deadlocks
	\end{itemize}

@@

	%...as well as subsequent P3 (kernel) grades
	...as well as subsequent kernel grades
	\begin{itemize}
		\item ``Does Landslide teach students lasting lessons about concurrency?''
	\end{itemize}
\end{frame}

@@

\begin{frame}{Example}
	\begin{center}
		\begin{tabular}{l}
			\texttt{if (\hilight{darkorange}{\_xbegin}() == SUCCESS) \{} \\ %\_XBEGIN\_STARTED) \{} \\
			%\texttt{} \\
			%\texttt{} \\
			\texttt{~~~~count++;} \\
			\texttt{~~~~\hilight{darkblue}{\_xend}();} \\
			\texttt{\} else \{}\\ %if (status != \_XABORT\_EXPLICIT) \{} \\
			\texttt{~~~~\hilight{darkorange}{mutex\_lock}(\&count\_lock);} \\
			\texttt{~~~~count++;} \\
			\texttt{~~~~\hilight{darkblue}{mutex\_unlock}(\&count\_lock);} \\
			\texttt{\}} \\
		\end{tabular}
	\end{center}
	% TODO: cite orig software TM, intel HTM
\end{frame}

@@

\begin{frame}{Example bug}
	\begin{center}
		\begin{tabular}{l|l}
			{\bf Thread 1} & {\bf Thread 2} \\
			\hline
			{\tt \hilight{darkorange}{\_xbegin}() == ABORTED} \\
			{\tt \hilight{darkorange}{mutex\_lock}(...);} \\
			{\tt int tmp = count;} \\
			{\em \hilight{pinkish}{<data-race PP>}} \\
				& {\tt \hilight{darkorange}{\_xbegin}() == SUCCESS} \\
				& {\tt count++;} \\
				& {\tt \hilight{darkblue}{\_xend}();} \\
			{\tt count = tmp + 1;} \\
			{\tt \hilight{darkblue}{mutex\_unlock}(...);} \\
		\end{tabular}
	\end{center}
\end{frame}

\begin{frame}{Example bug - Solution {\em [Dehesa-Azuara '16]}}
	\begin{center}
		\begin{tabular}{l}
			\texttt{if (\hilight{darkorange}{\_xbegin}() == SUCCESS) \{} \\ %\_XBEGIN\_STARTED) \{} \\
			%\texttt{} \\
			%\texttt{} \\
			\texttt{~~~~\hilight{pinkish}{if (mutex\_is\_locked(\&count\_lock))}} \\
			\texttt{~~~~~~~~\hilight{pinkish}{\_xabort();}} \\
			\texttt{~~~~count++;} \\
			\texttt{~~~~\hilight{darkblue}{\_xend}();} \\
			\texttt{\} else \{}\\ %if (status != \_XABORT\_EXPLICIT) \{} \\
			\texttt{~~~~\hilight{darkorange}{mutex\_lock}(\&count\_lock);} \\
			\texttt{~~~~count++;} \\
			\texttt{~~~~\hilight{darkblue}{mutex\_unlock}(\&count\_lock);} \\
			\texttt{\}} \\
		\end{tabular}
	\end{center}
\end{frame}

\begin{frame}{Example bug - Solution {\em [Dehesa-Azuara '16]}}
	\begin{center}
		\begin{tabular}{l|l}
			{\bf Thread 1} & {\bf Thread 2} \\
			\hline
			{\tt \hilight{darkorange}{\_xbegin}() == ABORTED} \\
			{\tt \hilight{darkorange}{mutex\_lock}(...);} \\
			{\tt int tmp = count;} \\
			{\em \hilight{pinkish}{<data-race PP>}} \\
				& {\tt \hilight{darkorange}{\_xbegin}() == SUCCESS} \\
				& {\tt \hilight{pinkish}{if (mutex\_is\_locked(...))}} \\
				& {\tt \hilight{pinkish}{~~~~\_xabort();}} \\
			{\tt count = tmp + 1;} \\
			{\tt \hilight{darkblue}{mutex\_unlock}(...);} \\
				& {\tt \hilight{darkorange}{mutex\_lock}(...);} \\
				& {\tt count++;} \\
				& {\tt \hilight{darkblue}{mutex\_unlock}(...);} \\
		\end{tabular}
	\end{center}
\end{frame}
